Contextuality and the Born rule in fuzzy quantum theories

### 2017-04-23 prop_letter.tex
prop_letter.tex Small things here and there...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Textbook quantum mechanics asserts that the probability of a measurement
can be computed from wave function of the system and the measuring
observable \emph{exactly} while real measurements must be finite precision.
Whether finite precision measurement in reality affect the idealized
quantum concept, e.g., contextuality~\cite{PhysRevLett.83.3751,Mermin1999,BarrettKent2004},
and the very foundation of quantum mechanics is still a problem. In
this letter, we model the finite precision quantum measurement by
quantum interval-valued probability measures (IVPMs) which provides
not only a uniform framework to understand Gleason's theorem~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}
and the Kochen-Specker theorem~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA}
but also some examples which cannot be explained by the Born rule~\cite{Born1983,peres1995quantum,544199,Jaeger2007}.

@@

Furthermore, if we express the observable~$\mathbf{O}$ as its spectrum
decomposition~$\sum_{i=1}^{N}\lambda_{i}P_{i}$, the expectation
value of $\mathbf{O}$ is given by~\cite{544199,Jaeger2007}
\begin{equation}
\expval{\mathbf{O}}_{\rho}=\Tr\left(\rho\mathbf{O}\right)=\sum_{i=1}^{N}\lambda_{i}\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)\ .\label{eq:quantum-expectation}
\end{equation}

@@

To formalize the idea of finite precision measurement, we next introduce
the notion of quantum IVPM~$\bar{\mu}$. The latter replaces the
infinitely precise interval $[0,1]$ by a finite collection $\mathscr{I}$
of intervals. Since the intervals $\imposs=\left[0,0\right]$ and
$\necess=\left[1,1\right]$ should be included in $\mathscr{I}$,
and arithmetic operations on intervals satisfy \begin{subequations}\label{eq:interval-operations}
\begin{eqnarray}
 &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}]\textrm{ and}\\
 &  & x[l,r]=\begin{cases}
[xl,xr] & \textrm{for }x\ge0\textrm{ ;}\\{}
[xr,xl] & \textrm{for }x\le0\textrm{ .}
\end{cases}
\end{eqnarray}
\end{subequations} $\bar{\mu}:\events\rightarrow\mathscr{I}$ is
defined by the axioms: 
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$. 
\item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
\item $\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)$. 
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
$P_{i}P_{j}=\delta_{ij}P_{i}$, we have 
\begin{equation}
\bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\textrm{ ,}\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{itemize}
condition that provides only a bound and not an equivalence as in
Eq.~\eqref{eq:QuantumProbability-Addition}. For example, if we consider
$\left\{ P_{1},P_{2}\right\} =\left\{ P,\mathbb{1}-P\right\} $ and
$\bar{\mu}\left(P\right)=\left[l,r\right]$, we have, when $r-l=\ell>0$,
\[
\bar{\mu}(\mathbb{1})=\left[1,1\right]\subsetneq\left[1-\ell,1+\ell\right]=\bar{\mu}\left(\mathbb{1}-P\right)+\bar{\mu}\left(P\right).
\]

@@

While mapping to $\mathscr{I}_{\infty}$ is one extreme case of quantum
IVPM, mapping to $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $
is another one, and corresponding to the Kochen-Specker theorem~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA}.
However, before diving into the Kochen-Specker theorem, it is better
to understand how to define expectation value of an observable with
quantum IVPMs.

@@

\yutsung{{Even if we have expectation value of }an observable, I
don't know how to reason Mermin square~(\ref{eq:MerminSquare}) directly
with quantum IVPM so far. The problem is the valuation function~$v$
is not an expectation value. Since the corresponding of Eq.~(\ref{eq:deterministic-product}),
$E\left[XY\right]=E\left[X\right]E\left[Y\right]$, doesn't hold for
expectation value except $X$ and $Y$ are uncorrelated or even independent~\cite{rohatgi2011introduction}.}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Textbook quantum mechanics asserts that the probability of a measurement
can be computed from wave function of the system and the measuring
observable \emph{exactly} while real measurements must be finite precision.
Whether finite precision measurement in reality affect the idealized
quantum concept, e.g., contextuality~\cite{PhysRevLett.83.3751,Mermin1999,BarrettKent2004},
and the very foundation of quantum mechanics is still a problem. In
this letter, we model the finite precision quantum measurement by
quantum interval-valued probability measures (IVPMs) which provides
not only a uniform framework to understand Gleason's theorem~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}
and the Kochen-Specker theorem~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA,Griffiths2003}
but also some examples which cannot be explained by the Born rule~\cite{Born1983,peres1995quantum,544199,Jaeger2007}.

@@

Furthermore, if we express the observable~$\mathbf{O}$ as its spectrum
decomposition~$\sum_{i=1}^{N}\lambda_{i}P_{i}$, the expectation
value of $\mathbf{O}$ is given by~\cite{peres1995quantum,544199,Jaeger2007}
\begin{equation}
\expval{\mathbf{O}}_{\rho}=\Tr\left(\rho\mathbf{O}\right)=\sum_{i=1}^{N}\lambda_{i}\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)\ .\label{eq:quantum-expectation}
\end{equation}

@@

To formalize the idea of finite precision measurement, we next introduce
the notion of quantum IVPM~$\bar{\mu}$. The latter replaces the
infinitely precise interval $[0,1]$ by a finite collection $\mathscr{I}$
of intervals. Since the intervals $\imposs=\left[0,0\right]$ and
$\necess=\left[1,1\right]$ should be included in $\mathscr{I}$,
and arithmetic operations on intervals satisfy \begin{subequations}\label{eq:interval-operations}
\begin{eqnarray}
 &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}]\textrm{ and}\\
 &  & x[l,r]=\begin{cases}
[xl,xr] & \textrm{for }x\ge0\textrm{ ;}\\{}
[xr,xl] & \textrm{for }x\le0\textrm{ .}
\end{cases}
\end{eqnarray}
\end{subequations} $\bar{\mu}:\events\rightarrow\mathscr{I}$ is
defined by the axioms: 
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$. 
\item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
\item $\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)$. 
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
$P_{i}P_{j}=\delta_{ij}P_{i}$, we have 
\begin{equation}
\bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{itemize}
condition that provides only a bound and not an equivalence as in
Eq.~\eqref{eq:QuantumProbability-Addition}. For example, if we consider
$\left\{ P_{1},P_{2}\right\} =\left\{ P,\mathbb{1}-P\right\} $ and
$\bar{\mu}\left(P\right)=\left[l,r\right]$, we have, when $r-l=\ell>0$,
\[
\bar{\mu}(\mathbb{1})=\left[1,1\right]\subsetneq\left[1-\ell,1+\ell\right]=\bar{\mu}\left(\mathbb{1}-P\right)+\bar{\mu}\left(P\right).
\]

@@

While mapping to $\mathscr{I}_{\infty}$ is one extreme case of quantum
IVPM, mapping to $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $
is another one, and corresponding to the Kochen-Specker theorem~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA,Griffiths2003}.
However, before diving into the Kochen-Specker theorem, it is better
to understand how to define expectation value of an observable with
quantum IVPMs.

@@

\yutsung{Even if we have expectation value of an observable, I don't
know how to reason Mermin square~(\ref{eq:MerminSquare}) directly
with quantum IVPM so far. The problem is the valuation function~$v$
is not an expectation value. Since $E\left[XY\right]=E\left[X\right]E\left[Y\right]$,
corresponding to Eq.~(\ref{eq:deterministic-product}), doesn't hold
for expectation value except $X$ and $Y$ are uncorrelated or even
independent~\cite{rohatgi2011introduction}.}
